Fitness tracking reveals task-specific associations between memory, mental health, and exercise

### 2021-10-16 paper/main.tex
updated bootstrap methods

\subsubsection*{Exploratory correlation analyses}
We used a bootstrap procedure to identify reliable correlations
between different memory-related, fitness-related, and
demographic-related variables.  For each of $N = 1000$ iterations, we
selected (with replacement) a sample of 113 participants to include.
This yielded, for each iteration, a sampled ``data matrix'' with one
row per sampled participant and one column for each measured variable.
When participants were sampled multiple times in a given iteration, as
was often the case, this matrix contained duplicate rows.  Next, we computed the Pearson's correlation
between each pair of columns.  This yielded, for each pair of columns,
a distribution of $N$ bootstrapped correlation coefficients.  If fewer
than $97.5\%$ of the coefficients for a given pair of columns had the
same sign, we excluded the pair from further analysis and considered
the expected correlation between those columns to be undefined.  If
$\geq 97.5\%$ of the coefficients for a given pair of columns had the
same sign (corresponding to a bootstrap-estimated two-tailed $p$
threshold of 0.05), we computed the expected correlation coefficient as:
\[
  \mathbb{E}_{i, j}\left[ r\right] = \tanh\left(\frac{1}{N} \sum_{n=1}^N
  \tanh^{-1}(\mathrm{corr}\left(m(i)_n, m(j)_n\right))\right),
\]
where $m(x)_n$ represents column $x$ of the bootstrapped data matrix
for iteration $n$, $\tanh$ is the hyperbolic tangent, and $\tanh^{-1}$
is the inverse hyperbolic tangent.

@@

We used the same bootstrap procedure described above (used in our
exploratory correlation analyses) to generate $N = 1000$ bootstrapped
data matrices whose rows reflected sampled participants and whose
columns reflected different measured variables.  We used a round-robin
imputation procedure to estimate the values of any missing
features~\citep{Buck60}.  We applied this imputation procedure
independently for each bootstrapped data matrix to prevent data
contamination across iterations.

Next, we fit a series of ridge regression models to the training data.
We grouped variables according to whether they were memory-related,
fitness-related, or demographic-related.  
We used variables from each group to predict held-out variables from
the same group-- e.g., using performance on all but one task to
predict performance on the held-out task.  We also examined
predictions across groups of variables-- e.g., using fitness and
demographic variables to predict memory-related variables.

@@

For each bootstrap iteration, we computed the root mean squared
deviation (RMSD) between the predicted and observed values in the
target features of the test dataset.  We defined this as the
\textit{observed} RMSD for that bootstrap iteration.  We also
estimated a null distribution of RMSD values by re-fitting each
iteration's data after randomly permuting the rows of the input data
matrix 
(thereby breaking any
meaningful associations between the input and output features).  Taken
together, this yielded a set of 1000 observed RMSD values and 1000
null RMSD values.  We assessed the statistical significance
($p$-values) of the observed RMSD values by computing the proportions
of null RMSD values that were less than the observed value.  We also
assessed the significance of the observed regression weights using
$t$-tests to compare the means of the observed versus null
distributions of weights, separately for each input and output feature.

>>>>

\subsubsection*{Exploratory correlation analyses}
We used a bootstrap procedure to identify reliable correlations
between different memory-related, fitness-related, and
demographic-related variables.  For each of $N = 10,000$ iterations, we
selected (with replacement) a sample of 113 participants to include.
This yielded, for each iteration, a sampled ``data matrix'' with one
row per sampled participant and one column for each measured variable.
When participants were sampled multiple times in a given iteration, as
was often the case, this matrix contained duplicate rows.  Next, we computed the Pearson's correlation
between each pair of columns.  This yielded, for each pair of columns,
a distribution of $N$ bootstrapped correlation coefficients.  If fewer
than $97.5\%$ of the coefficients for a given pair of columns had the
same sign, we excluded the pair from further analysis and considered
the expected correlation between those columns to be undefined.  If
$\geq 97.5\%$ of the coefficients for a given pair of columns had the
same sign (corresponding to a bootstrap-estimated two-tailed $p$
threshold of 0.05), we computed the expected correlation coefficient as:
\[
  \mathbb{E}_{i, j}\left[ r\right] = \tanh\left(\frac{1}{N} \sum_{n=1}^N
  \tanh^{-1}(\mathrm{corr}\left(m(i)_n, m(j)_n\right))\right),
\]
where $m(x)_n$ represents column $x$ of the bootstrapped data matrix
for iteration $n$, $\tanh$ is the hyperbolic tangent, and $\tanh^{-1}$
is the inverse hyperbolic tangent.

@@

We used the same bootstrap procedure described above (used in our
exploratory correlation analyses) to generate $N = 10,000$ bootstrapped
data matrices whose rows reflected sampled participants and whose
columns reflected different measured variables.  We used a round-robin
imputation procedure to estimate the values of any missing
features~\citep{Buck60}.  We applied this imputation procedure
independently for each bootstrapped data matrix to prevent data
contamination across iterations.

Next, we fit a series of ridge regression models to the training data.
We grouped variables according to whether they were memory-related,
fitness-related, or demographic-related.  We examined whether
fitness-related and demographic-related features could be used to predict
performance on one or more of the memory tasks.

@@

For each bootstrap iteration, we computed the root mean squared
deviation (RMSD) between the predicted and observed values in the
target features of the test dataset.  We defined this as the
\textit{observed} RMSD for that bootstrap iteration.  We also
estimated a null distribution of RMSD values by re-fitting each
iteration's data after randomly permuting the rows of the input data
matrix 
(thereby breaking any
meaningful associations between the input and output features).  Taken
together, this yielded a set of $10,000$ observed RMSD values and $10,000$
null RMSD values.  We assessed the statistical significance
($p$-values) of the observed RMSD values by computing the proportions
of null RMSD values that were less than the observed value.  We also
assessed the significance of the observed regression weights using
$t$-tests to compare the means of the observed versus null
distributions of weights, for each combination of features.
